---
title: "1000 Most Common Danish Words"
author: "Jemima"
date: "2023-11-25"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Gather Most Common Words

First, I am gathering the 1,000 most common Danish words. I chose this website to scrape from: <https://1000mostcommonwords.com/1000-most-common-danish-words/>

```{r echo = FALSE}
library(rvest)
library(dplyr)

link = "https://1000mostcommonwords.com/1000-most-common-danish-words/"
page = read_html(link)

    
Danish = page %>% html_nodes("tr+ tr td:nth-child(2)") %>% html_text()
# need to use html_nodes with a s
    
English = page %>% html_nodes("tr+ tr td~ td+ td") %>% html_text()
# html_node(webpage, css = 'selector')

Tusind_Ord <- data.frame(Danish, English)

Tusind_Ord %>% mutate(Number = row_number())

```

## Flash Cards

### Quiz Mode

Next, I will make a simple program that will flash each Danish word and prompt for its English translation. The program ends when I get something wrong.

```{r}
x = readline(prompt = "enter xth common words you want to play: ")
score = 0

while (TRUE) {
  n = sample(1:x, 1, replace = F)
  print(Tusind_Ord$Danish[n])
  answer = readline(prompt = "English meaning: ")
  if(answer == Tusind_Ord$English[n]){
    print("Correct")
    score = score + 1
  }
  else{
    print(paste0("Wrong, it is ", Tusind_Ord$English[n]))
    print(paste0("You got ", score, " words right."))
    break
  }
}
```

### Revision mode

The program will flash a Danish word. Upon command, the English word will be flashed after.

```{r}

while (n>0) {
  n = sample(1:1000, 1, replace = F)b
  print(Tusind_Ord$Danish[n])
  input = readline(prompt = "Press enter to flash the English Word (else press esc): ")
  if(input == ""){
    print(Tusind_Ord$English[n])
    second_input = readline(prompt = "Press enter to flash the next Word (else press esc): ")
      if(second_input != "" ){
        break
      }
  }
  else{
    break
  }
}
```
